\large{\bf Large effects and the infinitesimal model }

### 2023-12-02 review-responses.tex

\reply{
    We agree that \citet{turelli2017commentary} is clear about when the Gaussian emerges.
    It also seems quite clear to us that what we say in this paragraph agrees
    with \citet{turelli2017commentary}, who says in his Abstract that
    \begin{quote}
    This commentary distinguishes three nested approximations, referred to as “infinitesimal genetics,” “Gaussian descendants” and “Gaussian population,” each plausibly called “the infinitesimal model.” 
    \end{quote} 
    We think that the underlying issue here may be about motivation for the section.
    To make the motivation, and the point about the Gaussian population assumption more clear,
    we've added some explanatory text \revref.
    Also note that this section does \emph{not} make a Gaussian-population (or, non-Gaussian population) assumption;
    to make this more clear we've changed the phrase later in this paragraph
    from ``non-Gaussian model'' to ``model with non-Gaussian Mendelian sampling terms''.
}

\begin{point}{}
More generally, I found it a bit hard to follow the overall logical flow of section 4.1, especially once the authors launch into an exploration of the ``reproduction'' and ``noise'' terms (pages 9 and 10).
    It would be useful to clarify what the biological/intuitive meaning of the noise and reproduction terms at the outset. It would also be useful to state (at the beginning) what the main goal of these explorations is: (is the goal to identify when both terms ``have well-posed limits independent of the other'' as stated at \llname{limit_remark}), and also to summarise (at the end) to what extent one can do this in all generality.
\end{point}

\reply{
    Good suggestion; we've added more clarifying points;
    see \llname{clarification1}
    and \llname{clarification2}.
}

\begin{point}{}
    Last inequality above \revref: Is this the distribution of the largest of the $M$ alleles carried by an individual, conditioned on the trait value $Z$ of the individual (and this is independent of $Z$?), or is it just the largest of $M$ iid draws from a Cauchy distribution? If the former, then it'd be good to say this explicitly. If the latter, then this does not quite answer the question posed at the beginning of the section about how much information knowing the trait value of an individual gives about underlying allelic effects.
\end{point}

\reply{
    It is both --
    the calculation is, as the reviewer points out, about the largest of $M$ iid Cauchys,
    but those are the alleles carried by a parent in the initial generation.
    Furthermore, as explained at the beginning of this section \llname{crucial_step},
    this calculation about iid effects is indeed the key step
    (in \citet{Fisher1918} and \cite{barton2017infinitesimal})
    in describing the distribution of trait values in offspring.
    We've added an sentence that we think should remind the reader
    why we have in fact answered the question \revref.
}

@@

\reply{
    Good idea; we have done this \revref.
}

\subsubsection*{Minor comments:}

\begin{point}{}
    Introduction: Also worth citing Fisher, Bulmer etc. (the original references) when introducing the infinitesimal model?
\end{point}

\reply{
    Good idea, we have added some citations \revref.
    However, note that what exactly are the ``original references'' is not so clear:
    ``\textit{Indeed, its origins are lost in the mists of time (M. Bulmer, W.G. Hill, pers.  comm.).}''
    \citep[from][]{barton2017infinitesimal}
}

@@

\reply{
    These are good suggestions, however, we have encountered difficulties in implementing this,
    so we've settled for a note alerting the reader not to read too much into the absolute scale
    (see Figure~\ref{fig:sel_trait_distrns} caption).
}

\begin{point}{}
    Figure \ref{fig:seg_noise} caption: I find the phrasing ``(a,b) chosen to span the 5\% of midparents centered on the 10\% quantile of midparent value'' a bit confusing. Does this mean a and b are respectively the 7.5\% and 12.5\% quantile of the distribution of midparent trait values?
\end{point}

\reply{
    That's right; we've rephrased this confusing explanation.
}

\begin{point}{}
    Discussion: In the context of theory combining large and small-effect loci, \citet{chevin2008selective}: ``Selective Sweep at a Quantitative Trait Locus in the Presence of Background Genetic Variation'' and \citet{lande1983response}: ``The response to selection on major and minor mutations affecting a metrical trait'' may also be somewhat relevant.
\end{point}

\reply{
    Thanks! We've incorporated these \revref.
}

### 2023-12-02 writeup.tex
done with rev 1 points

Much of the work to 
understand the evolution of quantitative traits in sexual species
has focused on models in quantitative genetics
that describe evolution of trait values without explicitly representing underlying genetic variation,
but in a manner deriving from our understanding of how genetic variation works in practice.
Such models are therefore
interestingly intermediate between mechanistic models of genetics and purely phenomenological.
% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
% but tractable mathematical models of evolution for recombining species are elusive,
% at least with some results of when the model might be a good description of reality.
% For instance, a great deal of conceptual understanding and even quantitative predictions
% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
% despite the absence of a plausible model that would make this true (cite turelli and barton).
One concrete model of heritable trait evolution is the \emph{infinitesimal model},
which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
and a covariance independent of the parental trait values (but that depends on how they are related).
\citet{barton2017infinitesimal} give a history of the model
and a formal proof of conditions under which this would be a good approximation to reality,
which they extended to include dominance in \citet{barton2022infinitesimal}.

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits claimed to be genetically determined \revpoint{1}{14}
in fact only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race.
% \citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions.  We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8} when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.

@@

Unfortunately, once selection is included, \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.  We will content ourselves with presenting two examples of stabilizing selection, one already familiar to quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.

@@

The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, following the arguments in \cite{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits. Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above. A similar argument can be made for any $\alpha$-stable distribution, but the lack of explicit probability density functions for generic $\alpha \in (0,2)$ adds complexity to the argument without adding any insight.

@@

In particular, there is at least one locus at which knowing the parental trait gives one information about the effect size (and there are on the order of $\Oh\left(\frac{1}{\epsilon}\right)$ that give ``an $\epsilon$ of information''), and these large effect loci in turn give rise to correlations between
parental traits and the deviation of the offsprings' traits from the midparent
(i.e., the Mendelian sampling terms).
However, the first calculations above
suggest that for ``most'' loci,
this is not true,
and so perhaps a reasonable picture of Mendelian sampling terms
in an additive model with Cauchy-distributed effects
might be ``some large Mendelian alleles plus Gaussian noise''.
We now turn to simulation,
to investigate.

@@

A substantial challenge for the field is to develop tractable models of trait evolution
that include contributions of both ``large'' and ``small'' effect loci.
In practice, breeders have long embellished the ``animal model'' (essentially, the infinitesimal model)
to predict traits
by combining both effects of known large effect loci
and cumulative effects of many unknown loci as mediated by a relatedness matrix
\citep[e.g.,][]{fernando1989marker,teissier2018weighted,bernardo2014genomewide,rice2019evaluation}.
We can study any model with simulation,
but the way forwards to theoretical understanding seems murky.

>>>>

Much of the work to 
understand the evolution of quantitative traits in sexual species
has focused on models in quantitative genetics
that describe evolution of trait values without explicitly representing underlying genetic variation,
but in a manner deriving from our understanding of how genetic variation works in practice.
Such models are therefore
interestingly intermediate between mechanistic models of genetics and purely phenomenological.
% The action of natural selection on genetic variants in the absence of genetic variation is relatively well-understood (cite ewens),
% but tractable mathematical models of evolution for recombining species are elusive,
% at least with some results of when the model might be a good description of reality.
% For instance, a great deal of conceptual understanding and even quantitative predictions
% have been made assuming that the distribution of trait values in a population is Gaussian (cite Lande etcetera)
% despite the absence of a plausible model that would make this true (cite turelli and barton).
One concrete model of heritable trait evolution is the \emph{infinitesimal model},
which assumes a simple, explicit form for how the genetic component of offsprings' traits are determined by their parents:
they are Normally distributed, with the mean for each offspring equal to the average of the parents' values,
and a covariance independent of the parental trait values (but that depends on how they are related).
\citet{barton2017infinitesimal} give an overview of the complex history of the model, \revpoint{1}{12}
which is used in or adjacent to a good deal of theoretical work on quantitative genetics
\citep[e.g.,][]{fisher1918correlation,bulmer1971effect,lande1975maintenance}.
\citet{barton2017infinitesimal} focus on giving a formal proof of conditions
under which this would be a good approximation to reality,
which they extended to include dominance in \citet{barton2022infinitesimal}.

@@

The ways in which genetic variants affect an organisms' trait can be complex,
and interactions between alleles, between loci, and with the environment are often important in practice.
In fact, many traits that have been claimed to be mostly genetically determined \revpoint{1}{14}
only appear heritable through non-genetic means,
e.g., systematic inequalities in resource distribution partitioned by class or race
\citep[e.g.,][]{davenport1916hereditary}.
In this paper, we follow the literature in working with the relatively simple purely additive case
of a trait that can be decomposed into a sum across loci of the contributions of the alleles from each locus,
even omitting an independent effect of ``the environment''.
However, we emphasize that such models should not always be the default for understanding biological variation,
especially for human traits.

@@

Since $\varphi$ and $\Phi$ are characteristic functions, they are continuous and $\phi(0) = \Phi(0) = 1$, and thus both are non-zero in a neighbourhood of 0. In particular, $\psi(\theta) = \Log{\varphi}(\theta)$ and $\Psi(\theta) = \Log{\Phi(\theta)}$ exist in some neighbourhood of 0 as well\footnote{We use $\Log$ to denote the principal value of the complex logarithm, for which the imaginary part always lies in the interval $(-\pi,\pi]$.}.  From the above, we have
\[
	\Log{F^{(n)}(\varphi)(\theta)} = {\textstyle 2^{n} \psi\left(\frac{\theta}{2^{n}}\right)} 
		+ \sum_{k=0}^{n-1}  {\textstyle 2^{k} \Psi\left(\frac{\theta}{2^{k}}\right)}.
\]
We briefly explore what possible limiting values are valid characteristic functions. 
We refer to the first term on the left hand side as the \emph{reproduction term} and the remaining sum as the \emph{noise term}, and only consider the case \revpoint{2}{8}
when both have well-posed limits independent of the other. \llabel{limit_remark} % We don't need to limit to this caee.
Biologically,
the ``noise term'' is the contribution to variation in the population
generated by the Mendelian sampling terms (\ie, by recombination and mutation),
while the ``reproduction term'' is the contribution from variation in the initial generation.
\llabel{clarification2}

@@

Unfortunately, once selection is included, equation \eqref{eqn:MSB_CF} is less amenable to the sort of simple arguments above.  We will content ourselves with presenting two examples of stabilizing selection, one already familiar to quantitative genetics, the other a Cauchy solution obtained by trial-and-error, which illustrates how selection can stabilize an otherwise unviable solution.

@@

The crucial step in the derivation of the infinitesimal model in \cite{barton2017infinitesimal} is to show that, \llabel{crucial_step}
following the arguments in \citet{Fisher1918}, knowing an individual's trait value gives vanishingly little information about the contribution of any given locus as $M \to \infty$, which in turn allows one to show that the within-family Mendelian sampling term is independent of the parental traits. Here, we will consider how conditioning an individual's trait value affects the law of a single allelic contribution in the Cauchy setting described above. A similar argument can be made for any $\alpha$-stable distribution, but the lack of explicit probability density functions for generic $\alpha \in (0,2)$ adds complexity to the argument without adding any insight.

@@

To summarize, the key argument in \citet{Fisher1918} and \cite{barton2017infinitesimal}  \revpoint{1}{10}
requires that knowing the sum of parental effects does not give much information 
about any particular genotype;
we have shown here that this does not hold for Cauchy effect sizes
(at least when these are independent).
In particular, there is at least one locus at which knowing the parental trait gives one information about the effect size (and there are on the order of $\Oh\left(\frac{1}{\epsilon}\right)$ that give ``an $\epsilon$ of information''), and these large effect loci in turn give rise to correlations between
parental traits and the deviation of the offsprings' traits from the midparent
(i.e., the Mendelian sampling terms).
However, the first calculations above
suggest that for ``most'' loci,
this is not true,
and so perhaps a reasonable picture of Mendelian sampling terms
in an additive model with Cauchy-distributed effects
might be ``some large Mendelian alleles plus Gaussian noise''.
We now turn to simulation,
to investigate.

@@

A substantial challenge for the field is to develop tractable models of trait evolution
that include contributions of both ``large'' and ``small'' effect loci.
Various authors have taken the approach of separating contributions of large- and small-effect loci
(e.g., \citet{lande1983response} and \citet{chevin2008selective}). \revpoint{1}{20}
In practice, breeders have long embellished the ``animal model'' (essentially, the infinitesimal model)
to predict traits
by combining both effects of known large effect loci
and cumulative effects of many unknown loci as mediated by a relatedness matrix
\citep[e.g.,][]{fernando1989marker,teissier2018weighted,bernardo2014genomewide,rice2019evaluation}.
We can study any model with simulation,
but the way forwards to theoretical understanding seems murky.
